---
title: "conditionals visualizations"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, message=FALSE, echo=FALSE}
source("data-processing-functions.r")
source("webppl-model.R")
source("helpers.R")
require(plyr)
require(ggplot2)
require(reshape2)
require(gridExtra)

ALPHA <- 5
N_RUN <- 1
SAVE <- TRUE

TARGET_FN <- 'dirichlet-1'
```




```{r plot_helper-functions, echo=FALSE}
get_pc_pa_mus <- function(df){
  mu_pc <- ddply(df, "lt", summarise, grp.mean=mean(pc))
  mu_pa <- ddply(df, "lt", summarise, grp.mean=mean(pa))
  
  # mu_pc <- rename(mu_pc, c("grp.mean"="pc"))
  mu_pc$variable <- rep("pc",dim(mu_pc)[1])
  mu_pa$variable <- rep("pa",dim(mu_pa)[1])
  
  mus <- rbind(mu_pc,mu_pa)
  return(mus)
}

getMeanByLT <- function(df, col, which_lt){
  dataByLT <- df %>% filter(lt==which_lt)
  score <- mean(dataByLT[[col]])
  data  <- data.frame(score=score, lt=which_lt)
  return(data)
}

splitFacet <- function(x){
  facet_vars <- names(x$facet$params$facets)         # 1
  x$facet    <- ggplot2::ggplot()$facet              # 2
  datasets   <- split(x$data, x$data[facet_vars])    # 3
  new_plots  <- lapply(datasets,function(new_data) { # 4
    x$data <- new_data
    
    if(length(x$layers)==2){
      x$layers[[2]] <- NULL
    }
    x})
}  

savePlot <- function(bias, lt, alpha, n_run, width, height, fn_main='', lawn_theta=0.05){
  fn_id <- paste(get_filename(bias, "PL", alpha, n_run, lawn_theta))
  # target_dir <- paste("../plots/seed-", SEEDS[n_run], "/", sep="")
  target_dir <- get_target_dir("plots", TARGET_FN, n_run)
  target <- paste(target_dir, fn_id, fn_main, '.png', sep="")
  ggsave(target, width = 4, height=2.5)
  print(paste('saved plot to', target))
}
```

```{r plot_functions, message=FALSE, echo=FALSE}
plot_pc_pa_facet_lt <- function(bias, alpha, n_run, lawn_theta=0.05, folderName=""){
  warning("data for PL and LL needed!")
  # ---- facet for LL and PL ----- #
  df <- getSamples(bias, alpha, n_run, lawn_theta, folderName=folderName, which_lt = c("LL", "PL"))
  mus <- get_pc_pa_mus(df)
  
  df.long <-melt(df[,c("pc","pa","lt","cn")],id.vars=c("lt","cn"), measure.vars=c("pc","pa"))
  p <- ggplot(df.long, aes(x=value,color=variable,fill=variable)) +
       geom_density(alpha=.2) + facet_wrap(~lt, scales="free_y") + xlab('') +
       geom_vline(data=mus, aes(xintercept=grp.mean, color=variable), linetype="dashed") +
       theme(legend.title=element_blank(),text = element_text(size=8))
  p
  if(SAVE){savePlot(bias, "PL", alpha, n_run, 6, 3, "-pc-pa-facet-lt", lawn_theta)}

  return(p)
}

plot_pc_pa_facet_p <- function(bias, alpha, n_run, lawn_theta=0.05, folderName=""){
  warning("data for PL and LL needed!")
  # ----- facet for P(C) and P(A) ---- #
  df <- getSamples(bias, alpha, n_run, lawn_theta, folderName=folderName, which_lt = c("LL", "PL"))
  mus <- get_pc_pa_mus(df)
  
  df.long <-melt(df[,c("pc","pa","lt","cn")],id.vars=c("lt","cn"), measure.vars=c("pc","pa"))
  p <- ggplot(df.long, aes(x=value,color=lt)) +
       geom_density() + facet_wrap(~variable, scales="free") + xlab('') +
       theme(legend.title=element_blank(),text = element_text(size=8))
       geom_vline(data=mus, aes(xintercept=grp.mean, color=lt), linetype="dashed")

  p
  if(SAVE){savePlot(bias, lt, alpha, n_run, 4, 2.5, "-pc-pa-facet-p", lawn_theta)}
  return(p)
}

plotHist <- function(proposition, bias, lt, n_run, title, from, to, alpha=5, bins=20, folderName=""){
  samples <- read_data(bias, lt, n_run, alpha, lawn_theta, folderName)
  
  if(proposition=='A'){
    indices_in_interval <- getInterval(samples$pa, from, to)
    samplesInterval <- samples[indices_in_interval, ]
    p <- ggplot(samplesInterval, aes(x=pa)) + 
          geom_histogram(aes(y = ..count..), bins=bins) +
          theme(text = element_text(size=8)) +
          ggtitle(title) + 
          xlab(paste(from, '< P(A) <', to))
  }else if(proposition=='C'){
    indices_in_interval <- getInterval(samples$pc, from, to)
    samplesInterval <- samples[indices_in_interval, ]
    p <- ggplot(samplesInterval, aes(x=pc)) + 
        geom_histogram(aes(y = ..count..), bins=bins) +
        theme(text = element_text(size=8)) +
        ggtitle(title) + 
        xlab(paste(from, '< P(C) <', to))
  }else{
    stop('proposition must be either "A" or "C"')
  }
  ratio_in_interval <- dim(samplesInterval)[1] / dim(samples)[1]
  print(paste('ratio of samples inside interval:', ratio_in_interval))
  
  p
  if(SAVE){savePlot(bias, lt, alpha, n_run, 3, 3, paste("-hist-p", proposition, sep=""), lawn_theta)}

  return(p)
}

plotTables <- function(bias, lt, n_run, alpha, lawn_theta=0.05, folderName=""){
  data_samples <- read_data(bias, lt, n_run, alpha, lawn_theta, folderName)
  t0 <- data_samples$pca 
  t1 <- data_samples$pcna
  t2 <- data_samples$pnca
  t3 <- data_samples$pncna
  df <- melt(data.frame(t0=t0,t1,t2,t3))
  colnames(df) <- c("entry", "val")
  
  labs <- c("P(-C,-A)","P(-C,A)","P(C,-A)","P(C,A)")
  levels(df$entry) <- rev(labs)
  m <- aggregate(df[, "val"], list(df$entry), mean)
  df2 <- data.frame(entry=levels(df$entry), val=m$x)
  
  p <- ggplot(df, aes(x = val, color=entry)) +
        geom_density(show.legend = FALSE) +
        theme(text = element_text(size=8)) +
        scale_x_continuous(breaks=c(0,0.5,1)) +
        # ylab('density') + xlab('joint probability table entries') +
        geom_vline(aes(xintercept=val), data=df2, linetype="dashed", color="blue") +
        facet_wrap(~entry,scales="free_y")
  p
  if(SAVE){savePlot(bias, lt, alpha, n_run, 3, 3, "-tables", lawn_theta)}

  return(p)
}

plotCNs <- function(bias, lt, n_run, alpha, lawn_theta=0.05, folderName=""){
  data_samples <- read_data(bias, lt, n_run, alpha, lawn_theta, folderName)
  
  df <- as.data.frame(data_samples$cn)
  df <- count(df)
  df$freq <- round(x=df$freq / sum(df$freq),digits=2)
  colnames(df)
  
  p <- ggplot(df, aes(x=data_samples.cn, y=freq)) +
        geom_bar(stat="identity", show.legend = FALSE, fill="blue") +
        geom_text(aes(label=freq), vjust = -0.4, size=2.5) +
        xlab('') + ylab('') + 
        theme(text = element_text(size=10), axis.text.x = element_text(angle = 60, hjust = 1),
              legend.title=element_blank())
  p
  if(SAVE){savePlot(bias, lt, alpha, n_run, 5, 3.5, "-cns", lawn_theta)}

  return(p)
}

plotCNs_LLPL <- function(bias, alpha, n_run, folderName="", lawn_theta=0.05){
  df <- getSamples(bias, alpha, n_run, lawn_theta, folderName)
  ll_cn <- as.data.frame(table(filter(df, lt=='LL')$cn))
  ll_cn$lt <- rep("LL", nrow(ll_cn))
  ll_cn$Freq <- round(x=ll_cn$Freq / sum(ll_cn$Freq), digits=2)
  
  pl_cn <- as.data.frame(table(filter(df, lt=='PL')$cn))
  pl_cn$lt <- rep("PL", nrow(pl_cn))
  pl_cn$Freq <- round(x=pl_cn$Freq / sum(pl_cn$Freq),digits=2)
  data <- rbind(ll_cn,pl_cn)
  
  p <- ggplot(data, aes(x=Var1, y=Freq, fill=factor(lt))) +
        geom_bar(stat="identity", position="dodge", show.legend = FALSE) +
        geom_text(aes(label=Freq), vjust = -0.4, size=2.5) +
        xlab('') + ylab('') + 
        theme(text = element_text(size=10), axis.text.x = element_text(angle = 60, hjust = 1),
              legend.title=element_blank()) + 
        facet_wrap(~lt)
  
  p
  if(SAVE){savePlot(bias, "PL", alpha, n_run, 5, 3.5, "-facet-lt-cns", lawn_theta)}
  return(p)
}

plotTableDensityAllCNs <- function(bias, lt, n_run, alpha, lawn_theta=0.05, folderName=""){
  data_samples <- read_data(bias, lt, n_run, alpha, lawn_theta, folderName)

  keep <- c("pca","pcna","pnca","pncna", "cn")
  data <- data_samples[keep]
  df <- melt(data,id.vars="cn")
  p <- list(); i <- 1
  for(cn in CNs){
    print(cn)
    marginalsCN <- as.data.frame(prop.table(table(df$cn)))
    idx <- which(marginalsCN$Var1==cn)
    df2 <- df[which(df$cn==cn),]
  
    if(dim(df2)[1]==0){
      print(paste('cn', cn, 'has 0 probability - no output plot'))
    }else{
      marginal <- format(marginalsCN[idx, "Freq"], digits=2)
      evs <- aggregate(df2[,3], list(df2$variable), mean)
      colnames(evs) <- c("variable","value")
      p[[i]] <- ggplot(df2, aes(x=value, color=variable)) +
        geom_density() +
        geom_vline(data=evs, aes(xintercept=value), linetype="dashed") +
        xlab("") + ylab("") +  ggtitle(paste(cn, " (", marginal, ")", sep="")) +
        theme(legend.position='none', plot.title = element_text(size=8, hjust = 0.5),
              text = element_text(size=8)) +
        facet_wrap(~variable, ncol = 2, scales="free_y") 
      i <- i + 1
    }
  }
  g <- do.call(arrangeGrob, c(p, top=paste(bias,lt,sep="-")))  
  if(SAVE){
    fn <- get_filename(bias, lt, alpha, n_run, lawn_theta)
    target_dir <- get_target_dir("plots", TARGET_FN, n_run)
    target <- paste(target_dir, fn, "-table-densities.png", sep="")
    ggsave(file=target, g, width=5.5, height=7)
  }

  return(g)
}

plotConditionalProbsDensities <- function(which_densities, bias, lt, n_run, alpha, lawn_theta=0.05,
                                          folderName=""){
  df <- read_data(bias, lt, n_run, alpha, lawn_theta, folderName)
  
  # Density plots of conditional probabilities for PL and LL
  if(which_densities=='cp'){
    collist <- c('pAgivenC', 'pNCgivenNA')
    labelsEV <- c("E[P(A|C)]", "E[P(-C|-A)]")
    labelsDensity <- c("P(A|C)", "P(-C|-A)")
  }else{
    collist <- c('pCgivenA', 'pNAgivenNC')
    labelsEV <- c("E[P(C|A)]", "E[P(-A|-C)]")
    labelsDensity <- c("P(C|A)", "P(-A|-C)")
  }
  
  df2 <- df[,c(collist,"lt")]
  df2.long <- melt(df2,id.vars="lt")
  df2.long$variable <- mapvalues(df2.long$variable, from = collist, to = labelsDensity)
  data <- df2.long[which(df2.long$lt==lt),]
  
  target_dir <- paste("../plots/seed-", SEEDS[n_run], "/", sep="")
  fn <- get_filename(bias, lt, alpha, n_run, lawn_theta)
  target <- paste(target_dir, fn, "-cp-densities.png", sep="")
  
  p <- ggplot(data = data, aes(x=value)) +
        geom_density(aes(color=variable), alpha = 0.4, show.legend = FALSE) +
        xlab('')  + theme(text = element_text(size=8), legend.title=element_blank()) +
        facet_wrap( ~ variable, scales="free_y")
  
  ggsave(file=target, p, width=4, height=2.5)
  return(p)  
}

plotStrength <- function(biases, alpha, scoreProperty, n_run=1, folderName="", lawn_theta=0.05){
  data <- data.frame()
  for(bias in biases){
    df <- getSamples(bias, alpha, n_run, lawn_theta, folderName)
    pl <- getMeanByLT(df, scoreProperty, "PL")
    ll <- getMeanByLT(df, scoreProperty, "LL")
    
    df <- rbind(pl,ll)
    df$bias <- c(bias,bias)
    data <- rbind(data, df)
  }
  
  strengths <- data.frame()
  for(bias in biases){
    indices <- which(data$bias==bias)
    strength <- round(data[indices[1],"score"][1]/data[indices[2],"score"][1],3)
    df <- data.frame(bias=c(bias), strength=c(strength))
    strengths <- rbind(strengths, df)
  }
  
  p <- ggplot(data=data, aes(x=bias, y=score, fill=lt)) +
      geom_bar(stat="identity", position=position_dodge()) +
      theme(text = element_text(size=8), legend.title = element_blank()) +
      scale_x_discrete(labels=biases) + 
      ylab('Expected Value') +
      #geom_text(aes(label=LABELS[[scoreProperty]]), position = position_dodge(width = 1), vjust = -0.4, size=1.5) +
      annotate("text", x = seq(1,length(biases)), y = rep(0.05,length(biases)), label=strengths$strength , size=2) +
      ggtitle(LABELS[[scoreProperty]]) 
p
if(SAVE){savePlot(bias, "PL", alpha, n_run, 4, 2.5, paste("-", scoreProperty, "-strengths", sep=""), lawn_theta)}
return(p)
}

plotDistributions <- function(bias, alpha, which_p, n_run=1, folderName="", lawn_theta=0.05){
  df <- getSamples(bias, alpha, n_run, lawn_theta, folderName)
  p <- ggplot(df, aes(x=df[,which_p])) +
       geom_density(alpha=.2) + facet_wrap(~lt, scales="free_y") + xlab('') +
       # geom_vline(data=mus, aes(xintercept=grp.mean, color=variable), linetype="dashed") +
       theme(legend.title=element_blank(),text = element_text(size=8))
  p
  return(p)
}
```


```{r speaker-plots-functions, echo=FALSE, message=FALSE}
visualizeSpeakerExpectation <- function(condition, bias, alpha, n_run, lawn_theta=0.05){
  data_dir <- paste("../data/seed-", SEEDS[n_run], "/", sep="")
  fn <- get_filename(bias, "PL", alpha, n_run, lawn_theta)
  target <- paste(data_dir, fn, "-SE-", condition, ".rds", sep="")
  df <- readRDS(target)
      
  p <- ggplot(df, aes(x=reorder(Group.1,-x), y=x)) +
        geom_bar(stat="identity",fill='steelblue') +
        ylab('') +   xlab('') +
        geom_text(aes(label=x), vjust = -0.4, size=2) +
        theme(text = element_text(size=8),
            axis.text.x = element_text(angle = 60, hjust = 1))
  p    
  if(SAVE){savePlot(bias, lt, alpha, n_run, 3.5, 5, paste("-SE-", condition, sep=""), lawn_theta)}
  return(p)
}
```


Plot distributions P(C) and P(A) for the literal and the pragmatic listener respectively without assuming any particular bias for the utterance *If A, C*. 

```{r Uncertainty_A_C_1}
p <-plot_pc_pa_facet_lt("none", ALPHA, N_RUN, folderName = "dirichlet-1")
p
all_plots <- splitFacet(p)
all_plots[[1]]
all_plots[[2]]
```

Plot histograms for P<sub>PL</sub>(A) and P<sub>PL</sub>(C) in interval (0.1,0.9).

```{r Uncertainty_A_C_2}
plotHist("A", "none", "PL", N_RUN, "Pragmatic Listener", 0.1, 0.9, ALPHA)
plotHist("C", "none", "PL", N_RUN, "Pragmatic Listener", 0.1, 0.9, ALPHA)
```

Sample 1000 Bayes nets from the prior distribution: once with prior conditioned on P(C,A)>=theta, once with prior conditioned on P(C,¬A)>=theta and once with prior conditioned on P(C)>=theta, but P(C,A) and P(C,¬A) both < theta.
For each of these Bayes nets, consider P<sub>S</sub>.

```{r speaker}
# visualizeSpeakerExpectation("pca", "none", ALPHA, N_RUN)
# visualizeSpeakerExpectation("pcna", "none", ALPHA, N_RUN)
# visualizeSpeakerExpectation("pc", "none", ALPHA, N_RUN)
# visualizeSpeakerExpectation("pnc", "none", ALPHA, N_RUN)
```

Plot prior distribution for each entry of the probability tables.

```{r prior1, message=FALSE}
plotTables("none", "prior", 1, ALPHA)
```

Plot distribution for each entry of the probability tables for the literal listener.

```{r prior2, message=FALSE}
plotTables("none", "LL", N_RUN, ALPHA)
```
Plot Histograms of P(C)-samples from prior and from the literal listener. 

```{r indicatives-histograms}
plotHist("C", "none", "prior", N_RUN, "Pragmatic Listener", 0, 1, ALPHA)
plotHist("C", "none", "LL", N_RUN, "Pragmatic Listener", 0, 1, ALPHA)

```

Plot distributions over causal networks for the literal and the pragmatic listener.

```{r indicatives-cns}
plotCNs_LLPL("none", ALPHA, N_RUN)
```

# BISCUIT CONDITIONALS

```{r Biscuit_Conditionals}
p <- plot_pc_pa_facet_p("pizza", ALPHA, N_RUN)
p
all_plots <- splitFacet(p)
all_plots[[1]]
# all_plots[[2]]
```

# CONDITIONAL PERFECTION

```{r Conditional-Perfection}
plotCNs_LLPL("lawn", ALPHA, N_RUN)
plotConditionalProbsDensities("cp", "lawn", "PL", N_RUN, ALPHA)
plotConditionalProbsDensities("cp", "lawn", "LL", N_RUN, ALPHA)

plotStrength(c("lawn", "none"), ALPHA, "pNCgivenNA", N_RUN, folderName = "dirichlet-025")
plotStrength(c("lawn", "none"), ALPHA, "pAgivenC", N_RUN, folderName="dirichlet-025")

```

```{r Indicative-conditionals-context, message=FALSE}
plot_pc_pa_facet_lt("douven1", ALPHA, N_RUN)
plotCNs_LLPL("douven1", ALPHA, N_RUN)
# visualizeSpeakerExpectation("no-condition", "douven1", ALPHA, N_RUN)
# visualizeSpeakerExpectation("no-condition", "none", ALPHA, N_RUN)

```


## Discussion
Plot distribution for each table entry conditioned on each causal network.

```{r discussion, message=FALSE}
for(bias in c("none", "lawn", "douven1")){
  plotTableDensityAllCNs(bias, "PL", N_RUN, ALPHA)
  plotTableDensityAllCNs(bias, "prior", N_RUN, ALPHA)
  plotTableDensityAllCNs(bias, "LL", N_RUN, ALPHA)
}
```

